---
title: "Statistical Inference Part 1 - Simulation Exercise"
author: "Zorina Alliata"
date: "Tuesday, February 17, 2015"
output: pdf_document
---


<h1>Synopsis</h1>

The exponential distribution can be simulated in R with rexp(n, lambda) where n is the number of observations and lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda.
In these simulation exercises, we investigate the distribution of averages of 40 exponentials over a thousand observations (n=100), assuming the lambda = 0.2


<h1>Results</h1>

Create a thousand simulated averages of 40 exponentials, i.e. rexp(40,0.2)

```{r}
expdist <- rep(NA,1000)
for (i in 1:1000){
    expdist[i] <- mean(rexp(40,0.2))
}
```

The theoretical center of the mean is 1/lambda = 1/0.2: 5

```{r}
calcmean <- mean(expdist)
calcmean
```

The calculated mean is 5.044 and the theoretical mean is 5, so the variation is negligible.

The theoretical variance is ((1/0.2)^2)/40 : 0.625

```{r}
calcvar <- var(expdist)
calcvar
```

The calculated variance is 0.6333 and the theoretical variance is 0.625, so both distributions have similar variability.

We use the scale() function to plot the distribution and compare it to a normal distribution.

```{r, echo=FALSE}
expscale <- scale(expdist)
hist(expscale,probability=T, main="", ylim=c(0, 0.5))
lines(density(expscale))
# Compare with the standard normal distribution
curve(dnorm(x,0,1), -3, 3, col="red", add=T)
```

This is to be expected because of the CLT.
<p>

<b>Evaluating the coverage of the confidence interval</b>

 
A 95% confidence interval should contain, if we simulate a big number of them, the mean value for the exponential distribution ($1 / \lambda$) 95% of the time.
 

```{r}

set.seed(567)

lambda <- 0.2

# checks for each simulation if the mean is in the confidence interval

inconfint <- function(lambda) {

                ehats <- rexp(1000, lambda)

                se <- sd(ehats)/sqrt(1000)

                ll <- mean(ehats) - 1.96 * se

                ul <- mean(ehats) + 1.96 * se

                (ll < 1/lambda & ul > 1/lambda)

        }

# estimate the coverage in each round of simulations

coverage <- function(lambda) {

    covvals <- replicate(100, inconfint(lambda))

    mean(covvals)

}

# perform the simulation

simres <- replicate(100, coverage(lambda))

mean(simres)

```

 

The confidence interval contains the theoretical value 94.84% of the

time (close to the expected 95%).

